An Imaging Search for Post-MS Companions of Sirius B

### 2021-11-04 paper/paper.tex
revise contrast curves using 9.1 for apparant mag

\begin{abstract}
Discovery and characterization of post-MS planets is essential to study planetary system evolution and planet-star interaction during the most critical phases of stellar evolution. We present deep imaging of Sirius B, the closest and brightest white dwarf, in order to constrain post-MS planetary evolution in the Sirius system. We use Keck/NIRC2 in \Lp-band (\qty{3.776}{\micro\meter}) across three epochs in 2020 using the technique of angular differential imaging. Our observations are contrast limited out to \qty{1}{\au} and background limited beyond. The 5$\sigma$ detection limits from our best performing epoch are \numrange{17}{20.3} absolute magnitude. We consider multiple planetary formation pathways in the context of Sirius B's evolution to derive mass sensitivity limits using the ATMO2020 and Sonora Bobcat model grids. We achieve sub-Jupiter sensitivities and sub-AU separations, reaching \qtyrange{1.7}{3}{\jupitermass} at \qty{0.5}{\au} down to a sensitivity of \qtyrange{0.7}{1.5}{\jupitermass} at \textgreater\qty{1}{\au}. Consistent with previous results, we do not detect any companions around Sirius B. Our excellent detection limits, even without coronagraphy, demonstrate the capabilities for modern high-contrast imaging studies of nearby (\textless\qty{25}{\parsec}) white dwarfs.
\end{abstract}

@@

\section{Observations}\label{sec:obs}
\defcitealias{bonnet-bidaudADONISHighContrast2008a}{BB08}
\defcitealias{bondSiriusSystemIts2017}{B17}
\defcitealias{collaborationGaiaEarlyData2021a}{G21a}
\defcitealias{cummingsWhiteDwarfInitialFinal2018}{C18}
\begin{deluxetable}{cccc}
    \tablecaption{Parameters of the Sirius system adopted in this study.\label{tbl:system}}
    \tablehead{
        \colhead{parameter} & \colhead{value} & \colhead{unit} & \colhead{ref.}
    }
    \startdata
    $t_\mathrm{sys}$ & 225 & \unit{\mega\year} & \citetalias{bondSiriusSystemIts2017,cummingsWhiteDwarfInitialFinal2018} \\
    $\pi$ & \num{374.49\pm0.23} & mas & \citetalias{collaborationGaiaEarlyData2021a} \\
    $d$ & \num{2.6702\pm.0016} & pc & \citetalias{collaborationGaiaEarlyData2021a} \\
    $a$ & \num{20.016\pm0.014} & \unit{\au} & \citetalias{bondSiriusSystemIts2017} \\
    \cutinhead{Sirius A}
    $M_\star$ & \num{2.063\pm0.023} & \unit{\solarmass} & \citetalias{bondSiriusSystemIts2017} \\
    \cutinhead{Sirius B}
    $M_\star$ & \num{1.018\pm0.011} & \unit{\solarmass} & \citetalias{bondSiriusSystemIts2017} \\
    $M_\mathrm{MS}$ & \num{5.1\pm1.1} & \unit{\solarmass} & \citetalias{bondSiriusSystemIts2017,cummingsWhiteDwarfInitialFinal2018} \\
    $t_\mathrm{WD}$ & 126 & \unit{\mega\year} & \citetalias{bondSiriusSystemIts2017}\\
    $m^{\mathrm{L}^\prime}$ & \num{9.01\pm0.16} & & \citetalias{bonnet-bidaudADONISHighContrast2008a} \\
    $M^{\mathrm{L}^\prime}$ & \num{11.88\pm0.16} & & \citetalias{bonnet-bidaudADONISHighContrast2008a,collaborationGaiaEarlyData2021a} \\
    \enddata
\end{deluxetable}

@@

\begin{figure*}
    \centering
    % \epsscale{0.95}
    \plotone{residuals}
    \caption{The flat residuals of each epoch after PSF subtraction, derotating, and collapsing. The inner two FWHMs are masked out for each frame.}
    \label{fig:residuals}
\end{figure*}

\begin{figure*}
    \centering
    % \epsscale{0.95}
    \plotone{detection_maps}
    \caption{\textbf{top row:} The \textit{significance} maps for each epoch accounting for small-sample statistics \citep{mawetFundamentalLimitationsHigh2014}. Typically a critical value for detection is 5. \textbf{bottom row:} The STIM maps for each epoch calculated from each residual cube. The STIM probability has a typical cutoff threshold of 0.5 for significant detections. The inner two FWHMs are masked out for each map.}
    \label{fig:detection}
\end{figure*}

@@

\begin{figure*}
    \centering
    \plotone{contrast_curves}
    \caption{The contrast curves for the best performing algorithm from each epoch. The solid lines are the Gaussian 5$\sigma$ contrast curves and the dashed lines are the Student-t corrected curves. The expected upper limit for a dynamically stable orbit of \qty{1.5}{\au} is plotted as a vertical dashed line. The annular PCA curve cuts off because the innermost annulus was not able to detect a 100 S/N companion with 5$\sigma$ significance.}
    \label{fig:contrast}
\end{figure*}

\begin{figure*}
    \centering
    \plotone{mass_curves}
    \caption{Mass sensitivity curves derived from the 2020-11-21 epoch, which has the most sensitive contrast. The limits are calculated assuming an \Lp absolute magnitude for Sirius B of \num{11.9}. Both curves are truncated at \qty{1.5}{\au} due to the dynamical stability limit. \textbf{(a)} The absolute magnitudes are converted to masses using the ATMO2020 isochrone grid with non-equilibrium chemistry and weak convective mixing. The solid lines are the Gaussian 5$\sigma$ detection limits and the dashed lines are the Student-t corrected limits. The two ages represent the ages of two potential formation pathways, one of which is the system age (\qty{225}{\mega\year}), the other is the WD cooling age of Sirius B (\qty{126}{\mega\year}). The relative difference between the ages comes out to $\sim$\qty{0.3}{\jupitermass} at \qty{1}{\au}. \textbf{(b)} The absolute magnitudes are converted to masses using the Sonora Bobcat grid with solar metallicity (solid lines) and with +\qty{0.5}{dex} metallicity (dotted lines). For clarity, we only show the Gaussian contrast curves in this panel. The Sonora grid does not have atmospheric spectra for $T_\mathrm{eff}$ \textless \qty{200}{\kelvin}, which causes the cutoffs around \qty{1}{\jupitermass}. The relative difference due to the metallicity is $\sim$\qty{0.1}{\jupitermass}.}
    \label{fig:mass-curves}
\end{figure*}

@@

We show the contrast curves from the best performing algorithm for each reduction in \Cref{fig:contrast}. We determine the limiting sensitivities in terms of the planetary mass by first calculating the contrast-limited absolute magnitude using an \Lp-band magnitude for Sirius B of 9.01 (adapted from \citealp{bonnet-bidaudADONISHighContrast2008a}) and a distance modulus of \num{-2.87} \citep{collaborationGaiaEarlyData2021a}. We divide \Cref{fig:contrast} into two regimes: speckle-limited and background-limited. The speckle-limited regime exists from \qtyrange{0.2}{1}{\au} characterized by the increasing sensitivity with separation. Here we reach a median 5$\sigma$ detection limit of $\sim$19 magnitude (\Lp). This regime is constrained by the quality of the AO correction and the PSF subtraction method, mainly. The background-limited regime (\textgreater\qty{1}{\au}) is characterized by the flattening of the contrast curves, and is primarily constrained by the sky brightness. In this region, we reach 20.3 magnitude (\Lp) in the 2020-11-21 epoch. Our data is background-limited due to the relative brightness of the sky in \Lp (\qty{2.91}{mag\per sq. arcsec}\footnote{\url{https://www2.keck.hawaii.edu/inst/nirc2/sensitivity.html}}) compared to the pixel-to-pixel noise sources (e.g., read noise).

@@

\Cref{fig:mass-curves} shows our most sensitive contrast curve converted to mass limits under the different models. The first panel uses the ATMO2020 models to show how the choice of isochrone age leads to a $\sim$\qty{0.3}{\jupitermass} difference in the background-limited regime. The second panel uses the Sonora models to demonstrate the relatively minor effects ($\sim$\qty{0.1}{\jupitermass}) +\qty{0.5}{dex} has on the mass limits. We could not fully utilize the Sonora grid, though, because there are no atmospheric models tabulated for effective temperatures below \qty{200}{\kelvin}, which are precisely the models we need for the background-limited regimes. Overall, we constrain our detection limits to \qtyrange{1.7}{2.5}{\jupitermass} at \qty{0.5}{\au} (\ang{;;0.19}) in the speckle-limited regime and ultimately \qtyrange{0.6}{1.1}{\jupitermass} at \textgreater\qty{1}{\au} (\ang{;;0.38}) in the background-limited regime.

@@

Post-MS planetary evolution has historically been limited to theoretical work. Recently, though increasingly strong and diverse observational constraints have invigorated the field. We set out in this work to study the nearby Sirius system for post-MS companions around Sirius B. The Sirius system is one of the most well studied in history, with Sirius B being the target of companion searches from the visible to the IR. While it is highly unlikely a first-generation planet survived post-MS evolution, imaging efforts have gradually increased the sensitivity to second-generation planets. In this work, we presented high-contrast images of Sirius B in the near-IR. Our 5$\sigma$ sensitivity limits are the best that have been reached for Sirius B so far, reaching \num{20.3} \Lp absolute magnitude at \textgreater\qty{1}{\au}. We consider multiple planetary formation pathways yielding ages between \qtyrange{126}{225}{\mega\year} and explore the effects of enriched metallicity. We translate our sensitivity limits using the ATMO2020 and Sonora Bobcat grids to constrain our mass detection limits to \qtyrange{0.7}{1.5}{\jupitermass} at \textgreater\qty{1}{\au}. Our observations also show how the high precision of the parameters of the Sirius system directly benefits the sensitivity to planets. For example, the $\sim$4\% relative uncertainty of Sirius B's age translates to mass detection limit uncertainty below \qty{0.1}{\jupitermass}. Despite the high sensitivity of this study, we found no significant evidence for a companion around Sirius B, consistent with previous results.

>>>>

\begin{abstract}
Discovery and characterization of post-MS planets is essential to study planetary system evolution and planet-star interaction during the most critical phases of stellar evolution. We present deep imaging of Sirius B, the closest and brightest white dwarf, in order to constrain post-MS planetary evolution in the Sirius system. We use Keck/NIRC2 in \Lp-band (\qty{3.776}{\micro\meter}) across three epochs in 2020 using the technique of angular differential imaging. Our observations are contrast limited out to \qty{1}{\au} and background limited beyond. The 5$\sigma$ detection limits from our best performing epoch are \numrange{17}{20.4} absolute magnitude. We consider multiple planetary formation pathways in the context of Sirius B's evolution to derive mass sensitivity limits using the ATMO2020 and Sonora Bobcat model grids. We achieve sub-Jupiter sensitivities and sub-AU separations, reaching \qtyrange{1.6}{2.4}{\jupitermass} at \qty{0.5}{\au} down to a sensitivity of \qtyrange{0.7}{1.2}{\jupitermass} at \textgreater\qty{1}{\au}. Consistent with previous results, we do not detect any companions around Sirius B. Our excellent detection limits, even without coronagraphy, demonstrate the capabilities for modern high-contrast imaging studies of nearby (\textless\qty{25}{\parsec}) white dwarfs.
\end{abstract}

@@

\section{Observations}\label{sec:obs}
\defcitealias{bonnet-bidaudADONISHighContrast2008a}{BB08}
\defcitealias{bondSiriusSystemIts2017}{B17}
\defcitealias{collaborationGaiaEarlyData2021a}{G21a}
\defcitealias{cummingsWhiteDwarfInitialFinal2018}{C18}
\begin{deluxetable}{cccc}
    \tablecaption{Parameters of the Sirius system adopted in this study.\label{tbl:system}}
    \tablehead{
        \colhead{parameter} & \colhead{value} & \colhead{unit} & \colhead{ref.}
    }
    \startdata
    $t_\mathrm{sys}$ & 225 & \unit{\mega\year} & \citetalias{bondSiriusSystemIts2017,cummingsWhiteDwarfInitialFinal2018} \\
    $\pi$ & \num{374.49\pm0.23} & mas & \citetalias{collaborationGaiaEarlyData2021a} \\
    $d$ & \num{2.6702\pm.0016} & pc & \citetalias{collaborationGaiaEarlyData2021a} \\
    $a$ & \num{20.016\pm0.014} & \unit{\au} & \citetalias{bondSiriusSystemIts2017} \\
    \cutinhead{Sirius A}
    $M_\star$ & \num{2.063\pm0.023} & \unit{\solarmass} & \citetalias{bondSiriusSystemIts2017} \\
    \cutinhead{Sirius B}
    $M_\star$ & \num{1.018\pm0.011} & \unit{\solarmass} & \citetalias{bondSiriusSystemIts2017} \\
    $M_\mathrm{MS}$ & \num{5.1\pm1.1} & \unit{\solarmass} & \citetalias{bondSiriusSystemIts2017,cummingsWhiteDwarfInitialFinal2018} \\
    $t_\mathrm{WD}$ & 126 & \unit{\mega\year} & \citetalias{bondSiriusSystemIts2017}\\
    $m^{\mathrm{L}^\prime}$ & \num{9.1\pm0.2} & & \citetalias{bonnet-bidaudADONISHighContrast2008a} \\
    $M^{\mathrm{L}^\prime}$ & \num{11.97\pm0.2} & & \citetalias{bonnet-bidaudADONISHighContrast2008a,collaborationGaiaEarlyData2021a} \\
    \enddata
\end{deluxetable}

@@

\begin{figure*}
    \centering
    \epsscale{0.95}
    \plotone{residuals}
    \caption{The flat residuals of each epoch after PSF subtraction, derotating, and collapsing. The inner two FWHMs are masked out for each frame.}
    \label{fig:residuals}
\end{figure*}

\begin{figure*}
    \centering
    \plotone{detection_maps}
    \caption{\textbf{top row:} The \textit{significance} maps for each epoch accounting for small-sample statistics \citep{mawetFundamentalLimitationsHigh2014}. Typically a critical value for detection is 5. \textbf{bottom row:} The STIM maps for each epoch calculated from each residual cube. The STIM probability has a typical cutoff threshold of 0.5 for significant detections. The inner two FWHMs are masked out for each map.}
    \label{fig:detection}
\end{figure*}

@@

\begin{figure*}
    \centering
    \plotone{contrast_curves}
    \caption{The contrast curves for the best performing algorithm from each epoch. The solid lines are the Gaussian 5$\sigma$ contrast curves and the dashed lines are the Student-t corrected curves. The absolute magnitude are calculated using an absolute magnitude for Sirius B of 11.97. The expected upper limit for a dynamically stable orbit of \qty{1.5}{\au} is plotted as a vertical dashed line. The annular PCA curve cuts off because the innermost annulus was not able to detect a 100 S/N companion with 5$\sigma$ significance.}
    \label{fig:contrast}
\end{figure*}

\begin{figure*}
    \centering
    \plotone{mass_curves}
    \caption{Mass sensitivity curves derived from the 2020-11-21 epoch, which has the most sensitive contrast. The limits are calculated from the absolute magnitude derived in the contrast curves. Both curves are truncated at \qty{1.5}{\au} due to the dynamical stability limit. \textbf{(a)} The absolute magnitudes are converted to masses using the ATMO2020 isochrone grid with non-equilibrium chemistry and weak convective mixing. The solid lines are the Gaussian 5$\sigma$ detection limits and the dashed lines are the Student-t corrected limits. The two ages represent the ages of two potential formation pathways, one of which is the system age (\qty{225}{\mega\year}), the other is the WD cooling age of Sirius B (\qty{126}{\mega\year}). The relative difference between the ages comes out to $\sim$\qty{0.3}{\jupitermass} at \qty{1}{\au}. \textbf{(b)} The absolute magnitudes are converted to masses using the Sonora Bobcat grid with solar metallicity (solid lines) and with +\qty{0.5}{dex} metallicity (dotted lines). For clarity, we only show the Gaussian contrast curves in this panel. The Sonora grid does not have atmospheric spectra for $T_\mathrm{eff}$ \textless \qty{200}{\kelvin}, which causes the cutoffs around \qty{1}{\jupitermass}. The relative difference due to the metallicity is $\sim$\qty{0.1}{\jupitermass}.}
    \label{fig:mass-curves}
\end{figure*}

@@

We show the contrast curves from the best performing algorithm for each reduction in \Cref{fig:contrast}. We determine the limiting sensitivities in terms of the planetary mass by first calculating the contrast-limited absolute magnitude using an \Lp-band magnitude for Sirius B of 9.1 (adapted from Ks-band magnitude from \citealp{bonnet-bidaudADONISHighContrast2008a}) and a distance modulus of \num{-2.87} \citep{collaborationGaiaEarlyData2021a}. We divide \Cref{fig:contrast} into two regimes: speckle-limited and background-limited. The speckle-limited regime exists from \qtyrange{0.2}{1}{\au} characterized by the increasing sensitivity with separation. Here we reach a median 5$\sigma$ detection limit of $\sim$19 magnitude (\Lp). This regime is constrained by the quality of the AO correction and the PSF subtraction method, mainly. The background-limited regime (\textgreater\qty{1}{\au}) is characterized by the flattening of the contrast curves, and is primarily constrained by the sky brightness. In this region, we reach 20.4 magnitude (\Lp) in the 2020-11-21 epoch. Our data is background-limited due to the relative brightness of the sky in \Lp (\qty{2.91}{mag\per sq. arcsec}\footnote{\url{https://www2.keck.hawaii.edu/inst/nirc2/sensitivity.html}}) compared to the pixel-to-pixel noise sources (e.g., read noise).

@@

\Cref{fig:mass-curves} shows our most sensitive contrast curve converted to mass limits under the different models. The first panel uses the ATMO2020 models to show how the choice of isochrone age leads to a $\sim$\qty{0.3}{\jupitermass} difference in the background-limited regime. The second panel uses the Sonora models to demonstrate the relatively minor effects ($\sim$\qty{0.1}{\jupitermass}) the increased metallicity has on the mass limits. We could not fully utilize the Sonora grid, though, because there are no atmospheric models tabulated for effective temperatures below \qty{200}{\kelvin}, which are precisely the models we need for the background-limited regimes. Overall, we constrain our detection limits to \qtyrange{1.6}{2.4}{\jupitermass} at \qty{0.5}{\au} (\ang{;;0.19}) in the speckle-limited regime and ultimately \qtyrange{0.7}{1.1}{\jupitermass} at \textgreater\qty{1}{\au} (\ang{;;0.38}) in the background-limited regime.

@@

Post-MS planetary evolution has historically been limited to theoretical work. Recently, though increasingly strong and diverse observational constraints have invigorated the field. We set out in this work to study the nearby Sirius system for post-MS companions around Sirius B. The Sirius system is one of the most well studied in history, with Sirius B being the target of companion searches from the visible to the IR. While it is highly unlikely a first-generation planet survived post-MS evolution, imaging efforts have gradually increased the sensitivity to second-generation planets. In this work, we presented high-contrast images of Sirius B in the near-IR. Our 5$\sigma$ sensitivity limits are the best that have been reached for Sirius B so far, reaching \num{20.4} \Lp absolute magnitude at \textgreater\qty{1}{\au}. We consider multiple planetary formation pathways yielding ages between \qtyrange{126}{225}{\mega\year} and explore the effects of enriched metallicity. We translate our sensitivity limits using the ATMO2020 and Sonora Bobcat grids to constrain our mass detection limits to \qtyrange{0.7}{1.1}{\jupitermass} at \textgreater\qty{1}{\au}. Our observations also show how the high precision of the parameters of the Sirius system directly benefits the sensitivity to planets. For example, the $\sim$4\% relative uncertainty of Sirius B's age translates to mass detection limit uncertainty below \qty{0.1}{\jupitermass}. Despite the high sensitivity of this study, we found no significant evidence for a companion around Sirius B, consistent with previous results.
